Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-04-09 article.tex
Spell check

% Affiliations
\usepackage{authblk}
\title{
  Transmembrane helices are also 
  an overlooked source of major histocompatibility complex class II epitopes
  and evolutionary more conserved than expected by chance
}
\author[1]{Rich\`el J.C. Bilderbeek}
\author[2]{Frans Bianchi}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of 
Groningen, Groningen, The Netherlands}
\affil[2]{Frans' Institute, University of Groningen, Groningen, The Netherlands}

@@

Transmembrane helices (TMHs) in the human proteome
are an overrepresented source of epitopes on major 
histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
This finding may be a methodological artifact and using the proteome
of an actual pathogen nullifies this finding.
Additionally, it is unknown if TMH-II is also likelier to present
TMHs than expected by chance. 
Lastly, if the immune system if more tailored to detect polypeptides
derived from TMHs, can we confirm this by detecting the evolutionary
conservation of this behavior?
This study shows that MHC-I [also has/does not have] more
epitopes derived from a TMH for a pathogen proteome, when compared with
a host proteome.
Additionally, MHC-II binds to polypeptides derived from TMHs 
[less/equally/more] often than expected by chance.
Lastly, this study shows that in the mutation rate in TMHs of a pathogen
is [lower/as much/higher], hinting at 
evolutionary [conservation/neutrality/disruptive selection].
Our findings suggest that the immune system is [less/neutral/more]
vigilant to TMHs than expected by chance and this has left [a/no]
signal in the evolutionary history of the pathogen.

@@

{\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
adaptive immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2,
COVID-19

@@

For MHC-1, it was found that epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all human leukocyte antigen (HLA)-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs may be more 
conserved due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.

@@

\iffalse
	Transmembrane helices and strong MHC-II-binding peptides
	were predicted for a tuberculosis reference proteome 
	from \url{https://www.ebi.ac.uk/reference_proteomes}
	using the \verb;epitopeome; R package (\cite{epitopeome}).
	We picked the three MHC-II alleles that are most abundant 
	in the current human population, 
	which are DRB4*0101, HLA-DPA10103-DPB10402 
	\richel{
	  We decided for DPB1*0402, but there are multiple alleles matching.
	  I just picked the first one of the alleles matching, which are:
	  HLA-DPA10x-DPB10402, where x equals 103-110, 201-204, 301-303 or 401
	}
	and DQA1*0501/DQB1*0301 \richel{reference here}.
	The 5\% peptides with the lowest IC50 values were defined as binders.
	We then simply counted the number of amino acids that were present inside the
	cell, within the membrane or outside of the cell, as well as if it was part 
	of a strong MHC-II binding site, as shown in figue \ref{tab:results}.

	The \verb;epitopeome; R package (\cite{epitopeome}) binds 
	together the \verb;tmhmm; (\cite{tmhmm}) and \verb;netmhc2pan; 
	(\cite{netmhc2pan}) 
	R packages. \verb;tmhmm; provides an R interface to
	TMHMM (\cite{krogh2001predicting, sonnhammer1998hidden}), a tool
	to predict where membrane proteins' amino acids are located within the
	membrance.   
	\verb;netmhc2pan; provides an R interface to
	NetMHC2pan (\cite{jensen2018improved}), a tool
	to predict MHC-II binding to proteins.
\fi

@@

\iffalse
	The evolutionary conservation of TMHs was measured from a
	DNA alignment of multiple transmembrane proteins in multitiple
	species of the mycoplasma bacterial family
	\richel{how obtained exactly? how alignment done?}.

@@

We concluded that the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.
Because the full COVID-19 only has 4 TMHs \richel{check}, the percentages
of HMC-I epitopes being part of a TMH are likelier to be affected by
stochasticity. We chose to use COVID-19 regardless, as the thousands
of its time-dated genomic sequences are ideal for determining the 
evolutionary conservation of MHC-I detecting TMHs.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\input{ks/table_1.latex}

>>>>

% Affiliations
\usepackage{authblk}
\title{
  Transmembrane helices are also 
  an overlooked source of major histocompatibility complex class II epitopes
  and evolutionary more conserved than expected by chance
}
\author[1]{Richèl J.C. Bilderbeek}
\author[2]{Frans Bianchi}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of 
Groningen, Groningen, The Netherlands}
\affil[2]{Frans' Institute, University of Groningen, Groningen, The Netherlands}

@@

Transmembrane helices (TMHs) in the human proteome
are an overrepresented source of epitopes on major 
histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
It is unknown if this same overrepresentation
is also the case for the proteome of a pathogen.
Additionally, it is unknown if TMH-II is also likelier to present
TMHs than expected by chance. 
Lastly, if the immune system if more tailored to detect polypeptides
derived from TMHs, can we confirm this by detecting the evolutionary
conservation of this behavior?
This study shows that MHC-I [also has/does not have] more
epitopes derived from a TMH for a pathogen proteome, when compared with
a host proteome.
Additionally, MHC-II binds to polypeptides derived from TMHs 
[less/equally/more] often than expected by chance.
Lastly, this study shows that in the mutation rate in TMHs of a pathogen
is [lower/as much/higher], hinting at 
evolutionary [conservation/neutrality/disruptive selection].
Our findings suggest that the immune system is [less/neutral/more]
vigilant to TMHs than expected by chance and this has left [a/no]
signal in the evolutionary history of the pathogen.

@@

{\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
adaptive immunity, transmembrane domain, epitopes, T lymphocyte, 
MHC-1, MHC-I, MHC-2, MHC-II, COVID-19

@@

For MHC-I, it was found that epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all human leukocyte antigen (HLA)-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs may be more 
conserved due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.

@@

\iffalse
	Transmembrane helices and strong MHC-II-binding peptides
	were predicted for a tuberculosis reference proteome 
	from \url{https://www.ebi.ac.uk/reference_proteomes}
	using the \verb;epitopeome; R package (\cite{epitopeome}).
	We picked the three MHC-II alleles that are most abundant 
	in the current human population, 
	which are DRB4*0101, HLA-DPA10103-DPB10402 
	\richel{
	  We decided for DPB1*0402, but there are multiple alleles matching.
	  I just picked the first one of the alleles matching, which are:
	  HLA-DPA10x-DPB10402, where x equals 103-110, 201-204, 301-303 or 401
	}
	and DQA1*0501/DQB1*0301 \richel{reference here}.
	The 5\% peptides with the lowest IC50 values were defined as binders.
	We then simply counted the number of amino acids that were present inside the
	cell, within the membrane or outside of the cell, as well as if it was part 
	of a strong MHC-II binding site, as shown in figure \ref{tab:results}.

	The \verb;epitopeome; R package (\cite{epitopeome}) binds 
	together the \verb;tmhmm; (\cite{tmhmm}) and \verb;netmhc2pan; 
	(\cite{netmhc2pan}) 
	R packages. \verb;tmhmm; provides an R interface to
	TMHMM (\cite{krogh2001predicting, sonnhammer1998hidden}), a tool
	to predict where membrane proteins' amino acids are located within the
	membrane.   
	\verb;netmhc2pan; provides an R interface to
	NetMHC2pan (\cite{jensen2018improved}), a tool
	to predict MHC-II binding to proteins.
\fi

@@

\iffalse
	The evolutionary conservation of TMHs was measured from a
	DNA alignment of multiple transmembrane proteins in multiple
	species of the mycoplasma bacterial family
	\richel{how obtained exactly? how alignment done?}.

@@

We concluded that the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.
Because the full COVID-19 only has 4 TMHs \richel{check}, the percentages
of MHC-I epitopes being part of a TMH are likelier to be affected by
stochasticity. We chose to use COVID-19 regardless, as the thousands
of its time-dated genomic sequences are ideal for determining the 
evolutionary conservation of MHC-I detecting TMHs.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}
  \input{ks/table_1.latex}
  \caption{Percentage of MHC-I epitopes overlapping with transmembrane helix}
  \label{table:1}
\end{table}
